Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2020-06-13 manuscript/main.tex
Add citation for unpublished work on antigenic models
Also, removes unhelpful precedence statement.

The influenza community has long recognized the importance of incorporating HI phenotypes and other experimental measurements of viral phenotypes with existing forecasting methods into an extensible, open source framework that can be used by professional virologists in their vaccine design process \cite{Gandon:2016gz,Morris:2017ea,Lassig:2017hr}.
Although several distinct efforts have made progress in using HI phenotypes to evaluate the evolution of seasonal influenza \cite{Steinbruck:2014kq,Neher:2016hy}, published methods stop short of developing a complete forecasting framework wherein the evolutionary contribution of HI phenotypes can be compared and contrasted with new and existing fitness metrics.
However, unpublished work by {\L}uksza and L\"assig submitted to the WHO GISRS network incorporates antigenic phenotypes into fitness-based predictions \cite{luksza_personal_communication}.
Here, we provide the first such open source framework for forecasting the genetic composition of future seasonal influenza populations.
We apply this framework to HA sequence data shared via the GISAID EpiFlu database \cite{shu2017gisaid} and to HI and FRA titer data shared by WHO GISRS Collaborating Centers in London, Melbourne, Atlanta and Tokyo.
With this framework, we show that HI phenotypes enable more accurate long-term forecasts of H3N2 populations compared to previous metrics based on epitope mutations alone.
However, we also find that composite models based on phenotypic measures of antigenic drift and genotypic measures of functional constraint consistently outperform any fitness models based on individual genotypic or phenotypic metrics.

>>>>

The influenza community has long recognized the importance of incorporating HI phenotypes and other experimental measurements of viral phenotypes with existing forecasting methods into an extensible, open source framework that can be used by professional virologists in their vaccine design process \cite{Gandon:2016gz,Morris:2017ea,Lassig:2017hr}.
Although several distinct efforts have made progress in using HI phenotypes to evaluate the evolution of seasonal influenza \cite{Steinbruck:2014kq,Neher:2016hy}, published methods stop short of developing a complete forecasting framework wherein the evolutionary contribution of HI phenotypes can be compared and contrasted with new and existing fitness metrics.
However, unpublished work by {\L}uksza and L\"assig submitted to the WHO GISRS network incorporates antigenic phenotypes into fitness-based predictions \cite{Morris:2017ea,luksza_personal_communication}.
Here, we provide an open source framework for forecasting the genetic composition of future seasonal influenza populations using genotypic and phenotypic fitness estimates.
We apply this framework to HA sequence data shared via the GISAID EpiFlu database \cite{shu2017gisaid} and to HI and FRA titer data shared by WHO GISRS Collaborating Centers in London, Melbourne, Atlanta and Tokyo.
With this framework, we show that HI phenotypes enable more accurate long-term forecasts of H3N2 populations compared to previous metrics based on epitope mutations alone.
However, we also find that composite models based on phenotypic measures of antigenic drift and genotypic measures of functional constraint consistently outperform any fitness models based on individual genotypic or phenotypic metrics.
